---
title: "Introduction to the rcore package"
author: "Scott Chamberlain"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the rcore package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`rcore` - Client for the CORE API (<https://core.ac.uk/docs/>).
CORE (<https://core.ac.uk>) aggregates open access research
outputs from repositories and journals worldwide and make them
available to the public.

## Installation

The package can be installed directly from the CRAN repository:

```{r eval=FALSE}
install.packages("rcore")
```

Or you can install the development version:

```{r eval=FALSE}
devtools::install_github("ropensci/rcore")
```

Once the package is installed, simply run: 

```{r}
library("rcore")
```

## Obtaining an API key

The Core API requires an API key, and, as such, requires you to register for the key on the [Core Website](https://core.ac.uk/api-keys/register).  Once you register with your email address you will be sent an API key that looks a little like this:

`thisISyourAPIkeyITlooksLIKEaLONGstringOFletters`

## Using the API Key

Best practice is to set the API key as an environment variable for your system, and then call it in R using `Sys.getenv()`.  If you set the parameter in `.Renviron` it is permanently available to your R sessions.  There is a decent writeup of how to use the `.Renviron` file in [Colin Gillespie & Robin Lovelace's Efficient R Programming] (https://csgillespie.github.io/efficientR/3-3-r-startup.html#renviron).  Be aware that if you are using version control you do not want to commit the `.Renviron` file in your local directory.  Either edit your global `.Renviron` file, or make sure that `.Renviron` is added to your `.gitignore` file.

Within the `.Renviron` file you will add:

```
CORE_API=thisISyourAPIkeyITlooksLIKEaLONGstringOFletters
```

Once you've added the API key, restart your R session and test to make sure the key has been added using the command:

```{r, echo=FALSE}
Sys.getenv("CORE_API")
```

